### 2007-01-05 Documentation/part1.tex
*** empty log message ***

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. Some of these functions may not be directly related to user-defined
systematics, but may be mainly used in this context:
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSimBins}
{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}
SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]

A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with {\tt glbGetEminEmax}.
\end{function}
DOES THAT ALSO WORK WITH VARIABLE BINS? WHAT IS THE MEAING OF EMIN AND EMAX? [WW]
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. We recommend that you familiarize yourself with these functions at this point.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User-defined priors}
\index{norm}{User-defined priors}

@@

\index{norm}{Advanced tricks}
For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
\begin{description}
\item[Minimum $\boldsymbol{\chi^2}$ larger than threshold.] If a located degeneracy has a minimum $\chi^2$ larger than the corresponding confidence level threshold for the discussed quantity of interest, the degeneracy can be immediately ignored. This saves a lot of computation time.
\item[Locating degeneracies in more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge. For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt glbChiTheta13}, where $\stheta$ or $\deltacp$ is fixed at the best-fit value. The result can then be used as a starting point for {\tt glbChiAll} for the individual experiments with the systematics switched on again. 
\item[Forcing the minimizer into the targeted solution.]
In addition to switching off the systematics, it can be useful to reduce the input errors during some steps of the localization process in order to make the minimizer not to run away too much from the targeted solution.
The example on page~\pageref{ex:sgndeg} illustrates this with the
input error for $\ldm$: Since the guessed starting point might be
quite far away from the real degeneracy, the algorithm may in some cases
find the original solution instead of the degeneracy (which can
be immediately seen from the output vector). The input error
for $\ldm$ gives the algorithm a ``bias'' against the original solution.
However, note that the input error must not be too small in order
to avoid a significant contribution of the prior to the final $\chi^2$.
Alternatively, one could once again run {\tt glbChiAll} with the located 
minimum as {\tt in} vector, and $\ldm$ kept free.
\item[Finding degeneracies with multiple experiments.] For multiple experiments, it turns out to be useful to locate the degeneracies for individual experiments first. Then, all of the found degeneracies below the threshold can be tested for the combination of all experiments.
\end{description}
Finally, note that any degenerate solution below the confidence level threshold which can not be located makes the result appear better than it actually is. Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
\index{norm}{Degeneracies|)}

@@

\GLOBES\ can compute the probabilities in vacuum
with the following function:
\begin{function}
\GLBNS{VacuumProbability}
{\tt double glbVacuumProbability(int l, int m, int panti,double E, double L)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} and the baseline {\tt L} in vacuum. The parameter
{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos. 
\end{function}
In addition, the oscillation probabilities in matter can be obtained
with:
\begin{function}
\GLBNS{ProfileProbability}
{\tt double glbProfileProbability(int exp, int l, int m, int panti,double E)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in matter, where the matter density profile
is the one of experiment {\tt exp}. The parameter
{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
The matter density profile including baseline is the one from the last
evaluated experiment. 
\end{function}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\AEDL\ names}
\index{norm}{AEDL@\AEDL!names}
\label{sec:aedl_names}

Since in \AEDL\ rules, cross section, fluxes etc. carry a `name' by
which they can be refered to and in C they cary only an integer as
index it is sometimes difficult to figure out the correct correspondence.
Therefore the information about this correspondence obtained during
parsing is stored and can be accessed within C by these two functions.
\begin{function}
\GLBNS{NameToValue}
{\tt int glbNameToValue(int exp, const char* context, const char *name)}
Converts an \AEDL\ name given as argument {\tt name} into the corresponding
C index. The variable {\tt context} describes wether this name belongs
to a rule, channel, flux, energy or cross type environment. {\tt exp}
is the number of the experiment and can not be {\tt GLB\_ALL}. It returns
either the index in case of success or {\tt -1} in case the name was not
found. 
\end{function}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Event rates}
\index{norm}{Event rates}
\label{sec:event_rates}
One can also return event rates in \GLOBES , but this feature
requires some knowledge about the experiment definition. 
In fact, many of these functions are very advanced, which means
that the reader who wants to use them should be familiar with
\Secs~\ref{sec:channel} and~\Sec~\ref{sec:rules} of the \AEDL\ manual.

\index{norm}{Event rates}
\GLOBES\ currently supports rule-based and channel-based event rate functions, where the information is in written into a file
or returned in a list. The rule-based functions are:
\begin{function}
\GLBNS{ShowRuleRates}
{\tt int glbShowRuleRates(FILE* stream, int exp, int rule, int pos,
int effi, int bgi, int coeffi, int signal)} prints the binned rule rates as
list with energy and event rate to the file {\tt stream} (either an
open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
specific rule {\tt rule} have to be chosen, as well as the signal
or background rate {\tt signal} (either \GLBC{GLB\_SIG} or \GLBC{GLB\_BG}).
The position {\tt pos} refers to the component within the signal or 
background, and can also be {\tt GLB\_ALL}. The function may return
the rates with (\GLBC{GLB\_W\_COEFF}) or without (\GLBC{GLB\_WO\_COEFF})
overall efficiency coefficient, as it is specified by {\tt coeffi}. 
In addition, it may contain the post-smearing efficiencies (set
{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
post-smearing backgrounds (set
{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). The pre-smearing
efficiencies and backgrounds cannot be accessed at the rule level.
The return value
is $0$ if successful, and $-1$ if unsuccessful.
\end{function}
\begin{function}
\GLBNS{TotalRuleRate}
{\tt double glbTotalRuleRate(int exp, int rule, int pos,
int effi, int bgi, int coeffi, int signal)} returns the total rates with
the same parameters as {\tt glbShowRuleRates}.
\end{function}
The function {\tt glbTotalRuleRate} is especially useful if
one wants to draw bi-rate graphs with total event rates, or look
for the $(\deltacp,\theta_{13})$-degeneracy by the intersection of 
neutrino and antineutrino constant event rate curves.
%
In order to obtain information on the structure of the rules, 
a number of additional functions are provided:
\begin{function}
\GLBNS{GetNumberOfRules}
{\tt int glbGetNumberOfRules(int exp)} returns the number of
rules in experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetLengthOfRule}
{\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
the length of rule {\tt rule} in experiment {\tt exp}. The parameter
{\tt signal} can be either \GLBC{GLB\_SIG} for the number of signal
components or \GLBC{GLB\_BG} for the number of background components.
\end{function}
\begin{function}
\GLBNS{GetNormalizationInRule}
{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
returns the normalization (normalization or background center)
in rule {\tt rule} of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}
\begin{function}
\GLBNS{GetChannelInRule}
{\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
returns the channel number in rule {\tt rule} at the position {\tt pos}
of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}
\begin{function}
\GLBNS{GetCoefficientInRule}
{\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
returns the coefficient of the component {\tt pos} in rule {\tt rule} 
of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}

>>>>

One can also return the systematics function name by 
\GLBNS{GetChiFunction}
{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id)}
returns the name of the systematics $\chi^2$ function of a given experiment {\tt exp} and rule {\tt rule} for
systematics on or off as given by {\tt on\_off}. The user must 
ensure that the return string {\tt sys\_id} is large enough to hold the name of the function.
Similarly, a pointer to the systematics function can be returned either by name, or by experiment and rule selection:
\GLBNS{GetChiFunctionPtr}
{\tt glb\_chi\_function glbGetChiFunctionPtr(const char *name)}
returns a pointer to the systematics $\chi^2$ function with a specified name {\tt name}.
\GLBNS{GetChiFunctionPtrInExperiment}
{\tt glb\_chi\_function glbGetChiFunctionPtrInExperiment(int exp, int rule, int on\_off)}
returns a pointer to the systematics $\chi^2$ function of experiment {\tt exp} and rule {\tt rule}.
Systematics on or off can be accessed by {\tt on\_off}.

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. 
A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with \GLB{glbGetEminEmax}.
\end{function}
DOES THAT ALSO WORK WITH VARIABLE BINS? WHAT IS THE MEAING OF EMIN AND EMAX? [WW]
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. Additionally, the
functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.  We recommend that you familiarize yourself with these functions at this point.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{User-defined priors}
\index{norm}{User-defined priors}
\label{sec:userdefined}

@@

\index{norm}{Advanced tricks}
For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
\begin{description}
\item[Minimum $\boldsymbol{\chi^2}$ larger than threshold.] If a located degeneracy has a minimum $\chi^2$ larger than the corresponding confidence level threshold for the discussed quantity of interest, the degeneracy can be immediately ignored. This saves a lot of computation time.
\item[Locating degeneracies in more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge. For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt glbChiTheta13}, where $\stheta$ or $\deltacp$ is fixed at the best-fit value. The result can then be used as a starting point for {\tt glbChiAll} for the individual experiments with the systematics switched on again. 
\item[Forcing the minimizer into the targeted solution.]
In addition to switching off the systematics, it can be useful to reduce the input errors during some steps of the localization process in order to make the minimizer not to run away too much from the targeted solution.
The example on page~\pageref{ex:sgndeg} illustrates this with the
input error for $\ldm$: Since the guessed starting point might be
quite far away from the real degeneracy, the algorithm may in some cases
find the original solution instead of the degeneracy (which can
be immediately seen from the output vector). The input error
for $\ldm$ gives the algorithm a ``bias'' against the original solution.
However, note that the input error must not be too small in order
to avoid a significant contribution of the prior to the final $\chi^2$.
Alternatively, one could once again run {\tt glbChiAll} with the located 
minimum as {\tt in} vector, and $\ldm$ kept free.
\item[Finding degeneracies with multiple experiments.] For multiple experiments, it turns out to be useful to locate the degeneracies for individual experiments first. Then, all of the found degeneracies below the threshold can be tested for the combination of all experiments.
\item[Tracking algorithm.] If one scans a large portion of the parameter space with different input values, it is often useful to use the output from the previous minimization as educated guess for the next minimization. This works often very well if the degeneracy can be located in a part of the parameter space and the input parameter values change slowly enough (adiabatic transformation of the fit manifold). 
\item[Pre-scanning the parameter space.] In some cases, a very fast procedure can be a pre-scan 
of the relevant parameters using the very fast {\tt glbChiSys}. For example, for the intrinsic degeneracy, the location of the degeneracy in the $(\deltacp,\theta_{13})$ plane can easily and quickly found with {\tt glbChiSys} while keeping the other parameters fixed. Use this location then as a starting point for {\tt glbChiAll}.
\item[Manual scan of a subspace.] In some cases, the minimizer easily ends up in the wrong/unwanted solution, which is in most cases the already known best-fit solution. For example, for the octant degeneracy location, it is difficult to prevent the minimizer from running into the best-fit octant. In this case, one can scan one parameter (such as $\theta_{23}$) on a grid by choosing only valid values, and use {\tt glbChiNP} to marginalize over the other ones (while keeping \eg\ $\theta_{23}$ fixed). Therefore, the 6-parameter minimization is split into a 1-parameter manual minimization and a 5-parameter simultaneous minimization. Advantage: Runs with all \GLOBES\ versions. Disadvantage: Very slow. 
\item[Using Schwetz-priors (\GLOBES\ 3.0 and higher).] An elegant and fast method to prevent the minimizer from running into parts of the parameter space is to use user-defined priors and add a penalty as soon as a taboo zone is entered (\cf, \Sec~\ref{sec:userdefined}). This method was initially suggested by Thomas Schwetz.
\item[Using total rates.] In order to systematically locate all degeneracies (the eight-fold degeneracy), relatively reliable methods use the total appearance neutrino and antineutrino event rates of the experiment to determine educated guesses. The method goes as follows: Plot the curves with equal total rate in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane for both neutrinos and antineutrinos using the same rates as in the best-fit point. The curves will intersect at the intrinsic degeneracy (if it all). Plot the same curves for the same rates with $\mathrm{sgn}(\Delta m_{31}^2)$ flipped. Again you will find a maximum of two intersection points. Now do the same for $\pi/2 - \theta_{23}$ flipped, and for  $\mathrm{sgn}(\Delta m_{31}^2)$ combined with $\pi/2 - \theta_{23}$ flipped (mixed degeneracy). You will find at most two more intersection points in each case.
The results should look somewhat like this, where the best-fit point is not marked (the marks correspond to the points where the discrete degeneracies are located according to this specific algorithm):
\begin{center}
\includegraphics[width=0.95\textwidth]{ircurves}
\end{center}
 Altogether, there is a maximum of eight intersection points in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane, one of which is the best-fit point. These points can be using as starting points of the minimizer to locate the eight-fold degeneracy. Note that similar methods using the $\chi^2$ instead of the total rates have also been successfully used in the past. 
\end{description}
Finally, note that any degenerate solution below the confidence level threshold which cannot be located makes the result appear better than it actually is. Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
\index{norm}{Degeneracies|)}

@@

\GLOBES\ can compute the probabilities in vacuum
with the following function:
\begin{function}
\GLBNS{VacuumProbability}
{\tt double glbVacuumProbability(int l, int m, int panti,double E, double L)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} and the baseline {\tt L} in vacuum. The parameter
{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos. 
\end{function}
In addition, the oscillation probabilities in matter can be obtained
with:
\begin{function}
\GLBNS{ProfileProbability}
{\tt double glbProfileProbability(int exp, int l, int m, int panti,double E)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in matter, where the matter density profile
is the one of experiment {\tt exp}. The parameter
{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
The matter density profile including baseline is the one from the last
evaluated experiment. 
\end{function}
For a constant matter density profile, it is sufficient to specify the oscillation channel with
\begin{function}
\GLBNS{ConstantDensityProbability}
{\tt double glbConstantDensityProbability(int l, int m, int panti, double E, double L, double rho)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in constant matter, where the matter density profile
has the constant density {\tt rho} and the baseline is {\tt L}. The parameter
{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
\end{function}
If one in addition wants to use the low-pass filter feature in \GLOBES\ (see \Sec~\ref{sec:energy}), can can use
\begin{function}
\GLBNS{FilteredConstantDensityProbability}
{\tt double glbFilteredConstantDensityProbability(int exp, int l, int m, int panti, double E)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in constant matter, where the matter density and baseline, as well as the filter properties are taken from experiment {\tt exp}. The parameter {\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
\end{function}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Information from \AEDL\ files}
\label{sec:aedl_names}

In some cases, it is necessary to obtain information from the \AEDL\
files. Very basic information is supplied with
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSimBins}
{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}
SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]

In order to obtain information on the structure of the rules, 
a number of functions are provided:
\begin{function}
\GLBNS{GetNumberOfRules}
{\tt int glbGetNumberOfRules(int exp)} returns the number of
rules in experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetLengthOfRule}
{\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
the length of rule {\tt rule} in experiment {\tt exp}. The parameter
{\tt signal} can be either \GLBC{GLB\_SIG} for the number of signal
components or \GLBC{GLB\_BG} for the number of background components.
\end{function}
\begin{function}
\GLBNS{GetNormalizationInRule}
{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
returns the normalization (normalization or background center)
in rule {\tt rule} of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}
\begin{function}
\GLBNS{GetChannelInRule}
{\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
returns the channel number in rule {\tt rule} at the position {\tt pos}
of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}
\begin{function}
\GLBNS{GetCoefficientInRule}
{\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
returns the coefficient of the component {\tt pos} in rule {\tt rule} 
of the experiment {\tt exp}.
The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
(\GLBC{GLB\_BG}).
\end{function}
%
Similarly to the rules, one can find the number of channels of an experiment:
\begin{function}
\GLBNS{GetNumberOfChannels}
{\tt int glbGetNumberOfChannels(int exp)} returns the number of 
channels of experiment {\tt exp}.
\end{function}

\index{norm}{AEDL@\AEDL!names}
Since in \AEDL\ rules, cross section, fluxes etc. carry a `name' by
which they can be refered to and in C they cary only an integer as
index it is sometimes difficult to figure out the correct correspondence.
Therefore the information about this correspondence obtained during
parsing is stored and can be accessed within C by these two functions.
\begin{function}
\GLBNS{NameToValue}
{\tt int glbNameToValue(int exp, const char* context, const char *name)}
Converts an \AEDL\ name given as argument {\tt name} into the corresponding
C index. The variable {\tt context} describes wether this name belongs
to a rule, channel, flux, energy or cross type environment. {\tt exp}
is the number of the experiment and can not be {\tt GLB\_ALL}. It returns
either the index in case of success or {\tt -1} in case the name was not
found. 
\end{function}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Event rates}
\index{norm}{Event rates}
\label{sec:event_rates}
One can also return event rates in \GLOBES , but this feature
requires some knowledge about the experiment definition. 
In fact, many of these functions are very advanced, which means
that the reader who wants to use them should be familiar with
\Secs~\ref{sec:channel} and~\Sec~\ref{sec:rules} of the \AEDL\ manual.
Note that some of the event rate access has changed in \GLOBES\ 3.0,
because user-defined systematics requires very fast access which was
not possible with the old method.

A very simple function is for the total rate
\begin{function}
\GLBNS{TotalRuleRate}
{\tt double glbTotalRuleRate(int exp, int rule, int pos,
int effi, int bgi, int coeffi, int signal)} returns the total rates.
A specific experiment {\tt exp} and a 
specific rule {\tt rule} have to be chosen, as well as the signal
or background rate {\tt signal} (either \GLBC{GLB\_SIG} or \GLBC{GLB\_BG}).
The position {\tt pos} refers to the component within the signal or 
background, and can also be {\tt GLB\_ALL}. The function may return
the rates with (\GLBC{GLB\_W\_COEFF}) or without (\GLBC{GLB\_WO\_COEFF})
overall efficiency coefficient, as it is specified by {\tt coeffi}. 
In addition, it may contain the post-smearing efficiencies (set
{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
post-smearing backgrounds (set
{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). The pre-smearing
efficiencies and backgrounds cannot be accessed at the rule level.
\end{function}
The function {\tt glbTotalRuleRate} is especially useful if
one wants to draw bi-rate graphs with total event rates, or look
for the $(\deltacp,\theta_{13})$-degeneracy by the intersection of 
neutrino and antineutrino constant event rate curves.

There are several functions to directly print or save the event rate
information:
\begin{function}
\GLBNS{ShowRuleRates}
{\tt int glbShowRuleRates(FILE* stream, int exp, int rule, int pos,
int effi, int bgi, int coeffi, int signal)} prints the binned rule rates as
list with energy and event rate to the file {\tt stream} (either an
open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
specific rule {\tt rule} have to be chosen, as well as the signal
or background rate {\tt signal} (either \GLBC{GLB\_SIG} or \GLBC{GLB\_BG}).
The position {\tt pos} refers to the component within the signal or 
background, and can also be {\tt GLB\_ALL}. The function may return
the rates with (\GLBC{GLB\_W\_COEFF}) or without (\GLBC{GLB\_WO\_COEFF})
overall efficiency coefficient, as it is specified by {\tt coeffi}. 
In addition, it may contain the post-smearing efficiencies (set
{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
post-smearing backgrounds (set
{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). The pre-smearing
efficiencies and backgrounds cannot be accessed at the rule level.
The return value
is $0$ if successful, and $-1$ if unsuccessful.
\end{function}
\begin{function}
\GLBNS{ShowChannelRates}
{\tt int glbShowChannelRates(FILE *stream,
int exp, int channel, int smearing, int effi, int bgi)}
prints the binned channel rates as
list with energy and event rate to the file {\tt stream} (either an
open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
specific channel {\tt channel} have to be chosen.
The function may return
the rates before (\GLBC{GLB\_PRE}) or after (\GLBC{GLB\_POST})
the energy smearing, as it is specified by {\tt smearing}.
In addition, it may contain the pre- and post-smearing efficiencies (set
{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
pre- and post-smearing backgrounds (set
{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). Note that
the post-smearing efficiencies and backgrounds cannot be taken into
account if the rates are returned before the energy smearing.
The return value
is $0$ if successful, and $-1$ if unsuccessful.
\end{function}

\index{norm}{Event rates}
For rate vectors, \GLOBES\ currently supports rule-based and channel-based event rate functions, where typically pointers on the rate vectors are returned. The following pointer-based functions are currently supported:
\begin{function}
\GLBNS{GetChannelRatePtr}
{\tt double *glbGetChannelRatePtr(int exp, int ch, int pre\_post)} returns a 
pointer on the simulated rate vector of experiment {\tt exp} and channel {\tt ch}. Either
pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
rates can be accessed.
\end{function}
\begin{function}
\GLBNS{GetRuleRatePtr}
{\tt double *glbGetRuleRatePtr(int exp, int rule)} returns a 
pointer on the simulated rate vector of experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetSignalRatePtr}
{\tt double *glbGetSignalRatePtr(int exp, int rule)} returns a 
pointer on the simulated signal rate vector of experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetBGRatePtr}
{\tt double *glbGetBGRatePtr(int exp, int rule)} returns a 
pointer on the simulated background rate vector of experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetChannelFitRatePtr}
{\tt double *glbGetChannelFitRatePtr(int exp, int ch, int pre\_post)} returns a 
pointer on the fit rate vector of experiment {\tt exp} and channel {\tt ch}. Either
pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
rates can be accessed.
\end{function}
\begin{function}
\GLBNS{GetSignalFitRatePtr}
{\tt double *glbGetSignalFitRatePtr(int exp, int rule)} returns a 
pointer on the fit signal rate vector of experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetBGFitRatePtr}
{\tt double *glbGetBGFitRatePtr(int exp, int rule)} returns a 
pointer on the fit background rate vector of experiment {\tt exp} and rule {\tt rule}.
\end{function}
A simple example how to use these functions to print a rate vector is
\begin{quote}
{\tt
 int i; \\
   int n\_bins = glbGetNumberOfBins(EXP\_FAR); \\
  double *true\_rates\_N = glbGetRuleRatePtr(0, 0); \\
  printf("Simulated rates, experiment 0, rule 0: $\backslash$n"); \\
  for(i=0;i<n\_bins;i++) printf("\% g ",true\_rates\_N[i]); \\
  printf("$\backslash$n");
}
\end{quote}

% OLD FUNCTIONS:
% \begin{function}
% \GLBNS{GetChannelRates}
% {\tt int glbGetChannelRates(double** data,
% size\_t* length, int exp, int channel, int smearing)}
% writes the binned raw channel rates to the list {\tt data}
% and the length of this list to {\tt length}. 
% A specific experiment {\tt exp} and a 
% specific channel {\tt channel} have to be chosen.
% The function may return
% the rates before ({\tt smearing} is \GLBC{GLB\_PRE}) or after ({\tt smearing} is \GLBC{GLB\_POST})
% the energy smearing, where no user-defined data
% (pre-/post-smearing efficiencies or backgrounds) are taken into account.
% The return value is $-1$ if unsuccessful.
% \end{function}
% \begin{function}
% \GLBNS{GetUserData}
% {\tt int glbGetUserData(double** data,
% size\_t* length, int exp, int channel, int smearing, int bgeff)}
% writes the binned user-defined backgrounds or efficiencies 
% to the list {\tt data} and the length of this list to {\tt length}. 
% A specific experiment {\tt exp} and a 
% specific channel {\tt channel} have to be chosen.
% The function may return
% the pre- ({\tt smearing} is \GLBC{GLB\_PRE}) or post- ({\tt smearing} is \GLBC{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLBC{GLB\_BG}) 
% or efficiencies ({\tt bgeff} is \GLBC{GLB\_EFF}). 
% The return value is $-1$ if unsuccessful.
% \end{function}
% Since \GLOBES\ reserves the memory for the lists returned in these
% functions, which it allocates on an internal stack, one has to reset
% the stack at the end of the rates access with
% \begin{function}
% \GLBNS{ResetRateStack}
% {\tt void glbResetRateStack()} resets the rate stack used for the
% lists returned from {\tt glbGetChannelRates} or {\tt glbGetUserData}.
% \end{function}
% A code excerpt to show the channel rates may look like this:
% \begin{quote}
% {\tt
%  double* rates; \\
%   size\_t length; \\
%   glbGetChannelRates(\&rates,\&length,0,0,GLB\_PRE); \\
%   int i; \\
%   for(i=0;i<length;i++) printf("\%g $\backslash$n",rates[i]); \\
%   glbResetRateStack(); \\
% }
% \end{quote}
